Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to MIR downstream tasks

### 2023-06-15 abstract/abstract.tex
Updates from Overleaf

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we used self-supervised contrastive learning to uncover the underlying structure of Western tonal music by learning deep audio features to improve unsupervised music boundary detection. We applied deep convolutional neural networks and a triplet loss function to identify abstract and semantic musical elements without relying on their sonic qualities. This way, we replaced traditional acoustic features with deep audio embeddings, paving the way for high-level, sound-agnostic, and content-sensitive music representation for boundary detection.

Our cognitively-based approach for learning embeddings focuses on using full-resolution data and preserving high-level musical information. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy.

Preliminary results suggest that our musically-informed technique has significant potential for boundary detection tasks and possibly nearly all MIR downstream tasks that are not purely sonic-based. While it didn't outperform state-of-the-art results, it showed very competitive results and room for improvement. Consequently, music-motivated audio embeddings appear promising and potentially adaptable to other tasks constrained by data scarcity.

>>>>

In collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), we used self-supervised contrastive learning to uncover the underlying structure of Western tonal music by learning deep audio features to improve unsupervised music boundary detection. We applied deep convolutional neural networks and a triplet loss function to identify abstract and semantic high-level musical elements without relying on their sonic qualities. This way, we replaced traditional acoustic features with deep audio embeddings, paving the way for sound-agnostic and content-sensitive music representation for boundary detection.

Our cognitively-based approach for learning embeddings focuses on using full-resolution data and preserving high-level musical information which unfolds in the time domain. A key component in our methodology is the use of triplet networks, which provide an effective way of understanding and preserving the relative distances between different pieces of music. This model structure is integral to our work, allowing us to identify and maintain the nuanced relationships within our musical data. Drawing upon our domain expertise, we developed robust transformations to encode heuristic musical concepts that should remain constant. This novel approach aims to reconcile music and machine learning, enhancing machine listening models' efficacy through deep learning and triplet networks.

Preliminary results suggest that, while not outperforming state-of-the-art results, our musically-informed technique has significant potential for boundary detection tasks and, most likely, nearly all MIR downstream tasks that are not purely sonic-based.

While music-motivated audio embeddings appear promising, delivering competitive results with room for improvement and potentially adaptable to other tasks constrained by data scarcity, the question remains if such general-purpose audio representation can mimic human hearing.

### 2023-06-15 discussion/discussion.tex
Updates from Overleaf

@@

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to MIR downstream tasks. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

>>>>

Too little representation layer?

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/images/Mean group aggregation.png}
    \caption{Caption}
    \label{fig:enter-label}
\end{figure}

@@

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks to learn low-dimensional music latent representations with applications to music boundary detection. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

### 2023-06-15 evaluation/evaluation.tex
Updates from Overleaf

The developed model has been evaluated on the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS}. The "Structural Analysis of Large Amounts of Music Information" project aims to execute comprehensive structural analyses on a diverse spectrum of music. In contrast to traditional methods, SALAMI segments musical pieces into distinct sections, integrating perceptual, functional, and transcription analyses. Despite certain limitations, this novel approach offers a comprehensive and nuanced understanding of musical structure.

>>>>

The developed model has been evaluated on the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS}. The \textit{Structural Analysis of Large Amounts of Music Information} project aims to execute comprehensive structural analyses on a diverse spectrum of music. In contrast to traditional methods, SALAMI segments musical pieces into distinct sections, integrating perceptual, functional, and transcription analyses. Despite certain limitations, this novel approach offers a comprehensive and nuanced understanding of musical structure.

### 2023-06-15 figures/neural networks/activation.tex
Updates from Overleaf

\begin{figure}
    \centering
    \scalebox{0.8}{
% NEURAL NETWORK activation
% https://www.youtube.com/watch?v=aircAruvnKk&list=PLZHQObOWTQDNU6R1_67000Dx_ZCJB-3pi&index=1
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
\caption[Activation function \cite{tikz}.]{\small{The input layer with \textit{n} nodes is denoted as $a_i^{(0)}$ for $i \in \{1, \ldots, n\}$. The output layer with \textit{m} nodes is $a_j^{(1)}$ for $j \in \{1, \ldots, m\}$. Connections between these layers are governed by the weight matrix $\mathbf{W}^{(0)} \in \mathbb{R}^{m\times n}$ and bias vector $\mathbf{b}^{(0)} \in \mathbb{R}^{m}$, with elements $w_{j,i}$ and $b_j^{(0)}$ respectively. The activation function $\sigma$ operates element-wise on the sum of the product of the weight matrix, input vector, and bias vector.}}

>>>>

\begin{center}
\end{center}
\begin{figure}
    \centering
    \scalebox{1.0}{
% NEURAL NETWORK activation
\begin{tikzpicture}[x=2.7cm,y=1.6cm]
\centering
  \message{^^JNeural network activation}
  \def\NI{5} % number of nodes in input layers
  \def\NO{4} % number of nodes in output layers
  \def\yshift{0.4} % shift last node for dots

@@

\end{tikzpicture}
}
\caption[Activation function \cite{tikz}.]{\small{Input $a_i^{(0)}$ and output $a_j^{(1)}$ layers are connected by weight matrix $\mathbf{W}^{(0)}$ and bias vector $\mathbf{b}^{(0)}$, processed by activation function $\sigma$.}}

### 2023-06-15 future work/future work.tex
Updates from Overleaf

transforms
stems
different takes
k-fold cross validation
hyper-params: kernel size 0.005 times the sample rate
larger representation layer
Loss function margin. A margin of 0.2 can be a starting point, but it might not be the best value for every task. It's recommended to experiment with different values and choose the one that provides the best performance on your validation set.

>>>>

\begin{itemize}
  \item Transforms
  \item Stems
  \item Different takes
  \item k-fold cross validation
  \item Hyper-params: kernel size $0.005$ times the sample rate
  \item Larger representation layer
  \item Easy triplet mining \cite{XuanImprovedMining}
  \item Different boundary detection algorithms \cite{sf} was used because it was the default in MSAF \cite{MSAF}
\end{itemize}

Loss function margin. A margin of 0.2 can be a starting point, but it might not be the best value for every task. It's recommended to experiment with different values and choose the one that provides the best performance on your validation set.

### 2023-06-15 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music \footnote{In this thesis, "music" refers to the Western tradition, characterized by conventions, practices, and aesthetics that have developed primarily in Europe and North America. This usage does not deny the diversity and richness of other musical traditions worldwide but reflects the present study's focus on a specific cultural context.}, integral to human culture and history, has evolved alongside our societies since ancient times. Its study is complex due to its subjectivity, mutable ground truth, and various styles and cultural contexts.

XXXXXXX

>>>>

Music \footnote{Throughout this thesis, "music" refers to the Western tradition, defined by the conventions, practices, and aesthetics that primarily evolved in Europe and North America. This term's usage does not negate the diversity and richness of other musical traditions worldwide. Instead, it specifies this study's concentration on a particular cultural context.}, a cornerstone of human culture and history, has developed hand in hand with our societies since ancient times. However, the study of music is complex due to its inherent subjectivity, ever-changing ground truth, and the array of styles and cultural contexts it encompasses.

Prior research emphasizes the necessity of selecting suitable audio features to distinguish and label unique music segments for music segmentation tasks. The challenge lies in acquiring annotated data for these feature transformations, which can be time-consuming and expensive. In response, scholars have turned to unsupervised deep learning using readily available music audio \cite{unsupervisedlearndeepfeat, GrillMUSICANNOTATIONS}. This method has substantially improved the generalization capacity of machine-listening models in downstream applications. Among others, segmentation algorithms have seen significant enhancements and delivered state-of-the-art results \cite{Hernandez-Olivan2021MusicFeatures, Li2023MERT:Training}.

Beyond applications in music structure, the evolution of neural networks has empowered the creation of condensed numerical representations encapsulating crucial musical traits. This evolution has simplified the computational process for various task-specific elements, including but not limited to:

\begin{itemize}
\item Creating an audio embedding method that excels across numerous applications without the necessity for fine-tuning \cite{Turian2022HEAR:Representations}
\item Boosting the classification of environmental sounds \cite{Kim2020OneStrategies, CramerLOOKEMBEDDINGS}
\item Enhancing vocal-centric music tasks through cross-domain audio embeddings \cite{Kim2021LearningLoss}
\item Integrating task-specific and pre-trained features to optimize audio classification \cite{Hung2022Feature-informedClassification}
\item Developing a music similarity search engine for video production \cite{epidemic}
\item Improving Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}
\item Evaluating the effectiveness of speaker recognition via pre-trained model embeddings \cite{lightweight}
\item Embedding songs for artist identification to facilitate similarity comparisons \cite{contentmusicsimtriplet2020}
\item Addressing cross-modal text-to-music retrieval issues \cite{WonEmotionStories}
\item Enabling automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}
\end{itemize}

Furthermore, deep audio embeddings offer the benefit of transferability, establishing a foundation for multiple tasks. Compared to training a model from scratch, this approach saves computational resources and time \cite{}.

Given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise for Music Information Retrieval (MIR) downstream tasks, including boundary detection. However, it is yet to be determined whether a general-purpose audio representation can successfully emulate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have demonstrated generalizability across as many as fourteen music-understanding tasks \cite{Li2023MERT:Training}.

### 2023-06-15 introduction/Subsections/Objectives.tex
Updates from Overleaf

This study explores music's high-level tonal structure by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings applicable to subsequent MIR tasks.

The study aims to develop a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output -usually a fixed-size vector- produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a music piece that integrates high-level elements like melody contour, harmony relationships, and other comprehensive components such as the relationships between elements unfolding through time, arranging the embedding space to emphasize the clustering of these sophisticated musical content without considering sonic qualities or attributes.

The research will utilize MIR computational methodologies and musicological perspectives to explore musical compositions from various styles and genres.

This approach seeks to replicate human auditory capabilities in understanding and identifying abstract and semantic musical elements independent of their sonic qualities.

>>>>

Can deep audio embeddings be learned from reordering and scrambling sequences of musical information, and can audio effects be added to improve unsupervised music boundary detection?

This study explores the high-level tonal structure of music by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn and analyze music embeddings that can be applied to subsequent MIR tasks, with a particular focus on music boundary detection.

This study aims to develop and learn a unified numerical understanding (or embedding \footnote{A learned representation or embedding is a numerical output - usually a fixed-size vector - produced by a machine-learning model. Good representations should be versatile across various tasks and require limited supervision \cite{Turian2022HEAR:Representations}.}) for a piece of music that integrates the high-level relationships between musical elements as they unfold over time.

The research will employ MIR computational methodologies and musicological perspectives to explore musical compositions across various styles and genres.

This approach seeks to replicate human auditory capabilities by understanding and identifying abstract and semantic musical elements independent of their sonic qualities.

### 2023-06-15 introduction/Subsections/Previous work.tex
Updates from Overleaf

Prior research underscores the importance of choosing appropriate audio features to differentiate and label distinct music segments for music segmentation tasks. The hurdle lies in procuring annotated data for learning these feature transformations, which can be laborious and costly. To overcome this, scholars have turned to unsupervised deep learning, utilizing accessible music audio \cite{unsupervisedlearndeepfeat, GrillMUSICANNOTATIONS}. This approach has markedly enhanced the generalization ability of machine-listening models in downstream applications, with segmentation algorithms reaping benefits in the form of state-of-the-art outcomes \cite{Hernandez-Olivan2021MusicFeatures, SalamonDeepSegmentation, teacher, Li2023MERT:Training}.

Beyond music structure applications, modern advancements in neural networks have facilitated the creation of condensed numerical representations that capture critical musical characteristics. This progress has streamlined the calculation process for various task-specific elements. Such tasks include:

\begin{itemize}
\item Developing an audio embedding method that excels across numerous applications without the need for fine-tuning \cite{Turian2022HEAR:Representations}
\item Enhancing environmental sound classification \cite{Kim2020OneStrategies, CramerLOOKEMBEDDINGS}
\item Improving vocal-centric music tasks via cross-domain audio embeddings \cite{Kim2021LearningLoss}
\item Merging task-specific and pre-trained features to improve audio classification \cite{Hung2022Feature-informedClassification}
\item Designing a music similarity search engine for video producers \cite{epidemic}
\item Augmenting Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}
\item Investigating the efficacy of speaker recognition pre-trained model embeddings \cite{lightweight}
\item Embedding songs for similarity comparison via artist identification \cite{contentmusicsimtriplet2020}
\item Addressing the cross-modal text-to-music retrieval issue \cite{WonEmotionStories}
\item Facilitating automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}
\end{itemize}

Moreover, deep audio embeddings provide the advantage of transferability, serving as a basis for multiple tasks. This approach conserves computational resources and time compared to training a model from scratch \cite{HamelTransferSimilarity}.

Given their demonstrated effectiveness in the existing literature and their potential for transfer learning, deep audio embeddings hold substantial promise for MIR downstream tasks, and boundary detection is no exception. Nevertheless, it remains uncertain whether a general-purpose audio representation can effectively replicate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have managed to generalize across up to fourteen music understanding tasks \cite{Li2023MERT:Training}.

>>>>

Related research such as \cite{SalamonDeepSegmentation} and \cite{deepfeaturesegment} explore how the use of audio embeddings learned via unsupervised methods can enhance the performance of music segmentation algorithms leveraging the power of deep learning and data-driven feature learning to improve upon traditional, manually-engineered methods of music segmentation.

Both studies conclude that using deep learning techniques to derive audio embeddings improves the performance of music segmentation algorithms, thus providing a better alternative to traditional manual feature engineering methods.

### 2023-06-15 methods/subsections/deep learning.tex
Updates from Overleaf

Deep learning (DL), a subset of machine learning, employs deep neural networks to decipher hierarchical data representations, allowing models with multiple processing layers to abstract data across numerous levels \cite{LeCun2015DeepLearning}. DL models have significantly improved various domains due to their nonlinear modeling capabilities and scalability with large datasets.

@@

Deep Neural Networks (DNNs), computational models somewhat rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. DNNs evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns \cite{perceptron}.

>>>>

Deep learning (DL), a subset of machine learning, employs deep neural networks (DNNs) to decipher hierarchical data representations, allowing models with multiple processing layers to abstract data across numerous levels \cite{LeCun2015DeepLearning}. DL models have significantly improved various domains due to their nonlinear modeling capabilities and scalability with large datasets.

@@

DNNs, computational models somewhat rooted in biological neural systems, utilize layers of artificial neurons for tasks such as pattern recognition, classification, and regression. They evolve by adjusting connections between neurons during learning, a concept initiated in the mid-20th century but modernized with Frank Rosenblatt's single-layer perceptron in 1957, capable of learning linearly separable patterns \cite{perceptron}.

### 2023-06-15 methods/subsections/implementation.tex
Updates from Overleaf

In pattern recognition, models are designed to exhibit robustness against known invariances—transformations of input data—thereby ensuring consistent output. Interestingly, even unknown invariances not explicitly considered in the model's design can be accommodated due to the model's inherent learning capacity.

Our implementation is a slight tweak to the \textit{contrastive learning of lusical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a subset of ML called self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn independently from unlabeled data to create their supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, turning unsupervised problems into supervised ones by auto-generating labels. Benefits of SSL include reduced dependence on labeled data, which contributes to more robust and generalizable data representations.

As stated above, the methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as antagonistic pairs. The training process, which minimizes a triplet loss function, encourages the model to produce similar representations for positive and dissimilar ones for antagonistic pairs.

@@

\subsection{Deep Neural Network Architecture}

@@

Introduced by Bromley and LeCun \cite{Bromley1993SignatureNetwork}, Siamese Networks are deep learning architectures designed for tasks requiring comparison or similarity assessment between instances. The architecture comprises identical subnetworks that share the same parameters, improving memory usage and computational efficiency. Rather than learning specific features of individual classes, they focus on a similarity metric, making them ideal for limited or imbalanced datasets. Each subnetwork processes an input independently, combining the outputs to yield a similarity score. Training with shared weights enables the model to learn invariant input representation, improving comparison efficiency. This is accomplished through specialized loss functions such as contrastive or triplet loss, aiming to minimize the distance between similar inputs and maximize it for dissimilar ones.

The Triple Siamese Network extends this architecture, comparing three input instances instead of two. It aims to learn an embedding space wherein similar instances are closer and dissimilar ones are distant. The inputs, consisting of an anchor, a positive instance (similar to the anchor), and a negative instance (dissimilar to the anchor), are processed by identical subnetworks sharing the same parameters.

@@

This fully convolutional model reduces computational requirements and learns features at different scales through its multi-resolution architecture \ref{tab:samplecnn}.

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c}{The following layers are repeated depending on the strides and hidden parameters:} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected or representation layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnn}
\end{table}

@@

Previous research has utilized Convolutional Neural Networks (CNNs) with various features such as Mel-Scaled Log-magnitude Spectograms (MLS), Self-Similarity Matrices (SSM), and Self-Similarity Lag Matrices (SSLM) as inputs \cite{Hernandez-Olivan2021MusicFeatures}. However, features derived from raw audio may lack interpretability in some scenarios \cite{Schindler2020DeepTutorial}, and raw audio presents unique advantages despite being computationally demanding. It ensures the preservation of the original signal, potentially unearthing novel insights, and allows for direct feature extraction via advanced deep learning models \cite{learning}. Nevertheless, it comes with challenges such as high dimensionality due to high sample rates, requiring substantial computational resources. Moreover, it is sensitive to noise, adding complexity to the extraction process.

Time-domain processing naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although Mel-spectrogram-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Spectrograms capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music content may not always align with human perceptions \cite{Kim2020OneStrategies, Mesostructures2023}.

@@

The positive image in every triplet of input data must preserve its intelligible content when subjected to transformations, regardless of alterations in sonic qualities and artifacts. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements identical or remarkably close to the original track.

While we have experimented with helpful audio augmentation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0, Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our transformation chain using \textit{torchaudio}'s \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

\textbf{Amplitude effects}, we use gain to modify the signal's amplitude with a constant factor $g \in [-12, 0]$.

\textbf{Time-domain effects} such as speed change and stretch alter the signal's playback speed and duration without changing its pitch, respectively. The speed and stretch factors are $\alpha \in [0.9, 1.1]$ for speed change and $\beta \in [0.9, 1.1]$ for the stretch factor.

\textbf{Frequency-domain effects} adjust the frequency content or pitch of the signal, as in pitch shifting, which alters the pitch by $\Delta p \in [-1200, 1200]$.

\textbf{Nonlinear effects} such as overdrive add harmonic distortion using the parameter $d \in [0, 30]$ for overdrive

\textbf{Modulation effects}, including chorus and tremolo, use a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters and the tremolo's amplitude modulation frequency and depth $f_m \in [0.1, 100]$ for the tremolo's modulation frequency and $d_m \in [1, 101]$ for the tremolo's depth

\textbf{Reverberation effects} simulate physical space's acoustic reflections and reverberations, like reverb that applies an impulse response $h_R[n]$.

Finally, \textbf{Noise effects} add a noise signal $N[n]$ with a signal-to-noise ratio (SNR) in the range $[12, 100]$.

The positive signal $P[n]$ is generated according to the equation given below, where the symbols denote specific effects as indicated:

\begin{equation}\label{eq:positive_signal}
P[n] = A[n] \ast h_{G}[n] \ast h_{C}[n] \ast h_{D}[n] \ast h_{P_t}[n] \ast h_{R}[n] \ast h_{S}[n] \ast h_{T}[n] \ast h_{T_m}[n] + N[n]
\end{equation}

@@

@@

The computation goes as follows:

@@

\item We finally concatenate the shuffled audio chunks to generate the negative signal that will have similar production while the content is completely ruined:
\begin{equation}\label{eq:negative_signal}
N = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

Contrastive loss and triplet loss are two prominent loss functions employed in the fields of feature learning, similarity learning, and dimensionality reduction.

@@

Triplet loss function directs the learning process by endeavoring to minimize the distance between the anchor and positive instances and maximize the distance between the anchor and negative instances. Contrastive loss, on the other hand, works on pairs of instances \cite{supercontrast}. The inclusion of a margin parameter in the loss function guarantees a minimum separation between positive and negative instances in the embedding space. Given the nature of our training needs, we propose to train by using the former.

@@

\subsubsection{Training strategy}

>>>>

In pattern recognition, models are designed to exhibit robustness against known invariances—transformations of input data—thereby ensuring consistent output. Even unknown invariances not explicitly considered in the model's design can be accommodated due to the model's inherent learning capacity.

Our implementation is a slight tweak to the \textit{contrastive learning of musical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a subset of ML called self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn independently from unlabeled data to create their supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, turning unsupervised problems into supervised ones by auto-generating labels. Benefits of SSL include reduced dependence on labeled data, which contributes to more robust and generalizable data representations.

As stated above, the methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as antagonistic pairs. The training process encourages the model to produce similar representations for positive and dissimilar ones for antagonistic pairs.

@@

\subsection{Deep architecture design}

@@

Introduced by Bromley and LeCun \cite{Bromley1993SignatureNetwork}, Siamese Networks are deep learning architectures designed for tasks requiring comparison or similarity assessment between instances. The architecture comprises identical subnetworks that share the same parameters, improving memory usage and computational efficiency. Rather than learning specific features of individual classes, they focus on a similarity metric, making them ideal for imbalanced datasets. Each subnetwork processes an input independently, combining the outputs to yield a similarity score. Training with shared weights enables the model to learn invariant input representation, improving comparison efficiency. This is accomplished through specialized loss functions such as contrastive or triplet loss, aiming to minimize the distance between similar inputs and maximize it for dissimilar ones.

The Triple Siamese Network extends a binary contrastive architecture by comparing three input instances instead of two. It aims to learn an embedding space wherein similar instances are closer and dissimilar ones are distant. The inputs, consisting of an anchor, a positive instance (similar to the anchor), and a negative instance (dissimilar to the anchor), are processed by identical subnetworks sharing the same parameters.

@@

With 2.4 million trainable parameters, this fully convolutional model reduces computational requirements and learns features at different scales through its multi-resolution architecture \ref{tab:samplecnntable}.

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c}{The following layers are repeated depending on the strides and hidden parameters:} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected or representation layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnntable}
\end{table}

@@

Previous research has utilized CNNs with various features such as Mel-Scaled Log-magnitude Spectograms (MLS), Self-Similarity Matrices (SSM), and Self-Similarity Lag Matrices (SSLM) as inputs \cite{Hernandez-Olivan2021MusicFeatures}. However, features derived from raw audio may lack interpretability in some scenarios \cite{Schindler2020DeepTutorial}, and raw audio presents unique advantages despite being computationally demanding. It ensures the preservation of the original signal, potentially uncovering novel insights, and allows for direct feature extraction via advanced DL models \cite{learning}. Nevertheless, it comes with challenges, such as high dimensionality requiring substantial computational resources. Moreover, it is sensitive to noise, adding complexity to the extraction process.

Time-domain processing naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although audio feature-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Mel-Scaled Log-magnitude Spectograms, for example, capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music content may not always align with human perceptions \cite{Kim2020OneStrategies, Mesostructures2023}.

@@

The positive image in every triplet of input data must preserve its intelligible content when subjected to transformations, regardless of alterations in sonic qualities and processing artifacts. Maintaining the temporal structure and meaningfulness of the content allows it to present musical elements remarkably close to the original track.

While we have experimented with helpful audio augmentation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0, Kharitonov2020DataDomain}, the specific requirements of our experiments required the development of our transformation chain using \textit{torchaudio}'s \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

\textbf{Amplitude effects}: The signal's amplitude is modified by a constant factor using gain $g \in [-12, 0]$.

\textbf{Time-domain effects}: The signal's playback speed and duration are altered through speed change and stretching, preserving the perceptual musical relationships. The respective factors are $\alpha \in [0.9, 1.1]$ for speed change and $\beta \in [0.9, 1.1]$ for stretch.

\textbf{Frequency-domain effects}: The frequency content is adjusted through pitch-shifting, modifying the pitch by $\Delta p \in [-1200, 1200]$ cents.

\textbf{Nonlinear effects}: Nonlinear distortion is introduced via overdrive with a parameter $d \in [0, 30]$.

\textbf{Modulation effects}: Effects such as chorus and tremolo employ a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters, while the tremolo's amplitude modulation frequency and depth are controlled by $f_m \in [0.1, 100]$ and $d_m \in [1, 101]$, respectively.

\textbf{Reverberation effects}: These effects simulate a physical space's acoustic reflections and reverberations by applying an impulse response $h_R[n]$.

\textbf{Noise effects}: A noise signal $Noise[n]$ is added with a signal-to-noise ratio (SNR) in the range $[12, 100]$.

The positive signal $P[n]$ is generated according to the expression given below, where the symbols denote specific effects as indicated:

\begin{equation}\label{eq:positive_signal}
P[n] = A[n] \ast h_{G}[n] \ast h_{C}[n] \ast h_{D}[n] \ast h_{P_t}[n] \ast h_{R}[n] \ast h_{S}[n] \ast h_{T}[n] \ast h_{T_m}[n] + Noise[n]
\end{equation}

@@

Random parameter updates within hardcoded ranges generate unique audio $P[n]$ out of $A[n]$ for each run-through. Adding random white noise and varying SNR creates countless noisy waveform variations. Although possible combinations can be estimated by multiplying discrete parameter values, the presence of continuous parameters and randomness in noise generation effectively results in infinite unique audio versions.

@@

The computation for the negative signal $N[n]$ per every $A[n]$ goes as follows:

@@

\item We finally concatenate the shuffled audio chunks to generate the negative signal that will have similar production while the content is completely ruined:
\begin{equation}\label{eq:negative_signal}
N[n] = C_{\sigma(1)} \oplus C_{\sigma(2)} \oplus \ldots \oplus C_{\sigma(n)}
\end{equation}
\end{enumerate}

@@

@@

On the contrary of the also popular contrastive loss \cite{supercontrast}, the triplet loss function directs the learning process by minimizing the distance between the anchor and positive instances \textbf{and} maximizing the distance between the anchor and negative instances. Including a margin parameter in the loss function guarantees a minimum separation between positive and negative instances in the embedding space.

@@

\subsubsection{Hardware and training strategy}

The deep learning models were trained on a high-performance cloud computing setup hosted on \href{https://cloud.google.com/}{Google Cloud Platform}. The machine was of type \texttt{n1-standard-32}, equipped with an Intel Skylake processor and four NVIDIA Tesla T4 GPUs. The multiple GPUs allowed for efficient parallel processing, significantly reducing the training time for the deep learning models.
